Adding Rows to Lumibatch Phenotdata dataframes 
=======================================================


```{r}



```


## Setwd
```{r setwd, eval=TRUE, tidy=TRUE}
setwd("/home/daniel/Documents/A_Year_1_PhD/GAP_Data/2015_01_29_GAP_Final_Limma_analysis")


getwd()
```

## Load Libs
```{r load_libs, tidy=TRUE}

library(lumi)
library(limma)
library(xtable)

```


## Load eset
```{r load_eset, tidy=TRUE}
load("GAP_DL_FINAL.eset_final.RData")
str(eset_final)

loadcannabisdata <- read.csv("GAP Gene-Ex manifest_added info_22-01-14.csv", head=TRUE, sep=",",na.string=c("NA","Unknown"),as.is=T)
load_age_and_dup_data <- read.csv("Enhanced_Age_DUP_Conrad_05-03-14.csv", head=TRUE, sep=",",as.is=T, na.string=c("NA","Unknown") )
load_meds <- read.csv  ("GAP_Medication_Daniel_2014.csv", head=TRUE,sep=";",as.is=T, na.string=c("NA","Unknown"))
load_master <- read.csv  ("GAP_Relevant_data_from_masterspread_sheet_with_study_ID3.csv", head=TRUE,sep=",",as.is=T, na.string=c("."))
```


##Check that Ethinicity (in cannabis data) and age  dup are numeric.
```{r}

load_age_and_dup_data$DUP_enhanced <- as.numeric(load_age_and_dup_data$DUP_enhanced)

```

## Create neweset. which is free of gender failed samlples, unknowns (case control status), and duplicates. 
```{r, tidy=TRUE}

pheno_data <- pData(eset_final)


samples_gendercheck_pass <- subset(pheno_data, pheno_data$XIST_PRKY_Combined_Gender_Fail=="PASS",useNA="ifany") # Remove gender failed samples

samples_gendercheck_duplicate_pass  <-samples_gendercheck_pass[duplicated(samples_gendercheck_pass$Study_ID)==F,] #Remove duplicates

samples_gendercheck_duplicate_unknown_pass  <-samples_gendercheck_duplicate_pass[samples_gendercheck_duplicate_pass$PHENOTYPE!="UNKNOWN",] #Remove unknown phenotype

samples_only_baseline_pass  <-samples_gendercheck_duplicate_unknown_pass[samples_gendercheck_duplicate_unknown_pass$Groups_orginal=="BASELINE",] 

GOOD_SAMPLES <- samples_only_baseline_pass$sampleID

neweset <- eset_final[,GOOD_SAMPLES]
```

##create New Pheno Data and add sample order!!!
```{r}

new_pheno_data <- pData(neweset)
new_pheno_data$sample_order <- 1:dim(new_pheno_data)[1] ##add sample order column
str(new_pheno_data)

dim(new_pheno_data [new_pheno_data$PHENOTYPE =="CONTROL",]) ## 192 Cases, 155 CONTROL
str(new_pheno_data)

colnames(new_pheno_data)[colnames(new_pheno_data)=="Study_ID"] <- "study_ID"
str(new_pheno_data)


```

##Tidy up a bit
```{r}
ethnicity_sampleID<- loadcannabisdata [,c("african","Indian","sampleID","groups")]

age_sampleID<- load_age_and_dup_data [,c("Age_enhanced","Groups","study_ID","sampleID")]

med_study_ID <- load_meds [,c("yes_no","total_cp","study_ID","case_control_status")]

master_dob_study_ID <- load_master [,c("dob","study_ID")]

master_ethnicity_self_study_ID <- load_master [,c("study_ID","ethnicity_selfreportfolders")]

```


##Ehnicity
```{r}
loadcannabisdata1 <-  ethnicity_sampleID[ethnicity_sampleID$groups=="baseline" ,]
ga_african <- loadcannabisdata1[complete.cases(loadcannabisdata1$african),] ##Remove NA  
  
  
##ga_african$african ##check that NA are removed
##ga_african$Indian ##check that NA are removed


```


##Age
```{r}
just_baseline_age <-  age_sampleID[age_sampleID$Groups=="baseline" ,]
unknown_removed_age <- just_baseline_age [complete.cases(just_baseline_age$Age_enhanced),]

```

##Medication
```{r}

load_meds_d_removed  <-med_study_ID[duplicated(med_study_ID$study_ID)==F,] #Remove duplicates

```

##Master
```{r}

# Master_DOB
str(master_dob_study_ID)
master_dob_study_ID
master_dob_no_NA  <- master_dob_study_ID[master_dob_study_ID$dob!="NA",] 

#Master_ethnicity_self
str(master_ethnicity_self_study_ID)
master_ethnicity_self_no_NA  <- master_ethnicity_self_study_ID[master_ethnicity_self_study_ID$ethnicity_selfreportfolders!="NA",] 
master_ethnicity_self_no_NA

```




##Merge everything
```{r}

# Conrads Data
merged_age_sex  <- merge (unknown_removed_age,new_pheno_data, by="sampleID")
merged_age_sex_ethnicity  <- merge (merged_age_sex,ga_african, by="sampleID")
merged_age_sex_ethnicity_meds  <- merge (merged_age_sex_ethnicity,load_meds_d_removed, by="study_ID")

str(new_pheno_data)
str(merged_age_sex)
str(merged_age_sex_ethnicity_meds)
str(merged_age_sex_ethnicity_meds_duplicate_pass)
table(merged_age_sex_ethnicity_meds$SEX)


# Master Data
merged_dob_sex  <- merge (master_dob_no_NA,new_pheno_data, by="study_ID")
merged_dob_sex_ethnicity  <- merge (merged_dob_sex, master_ethnicity_self_no_NA, by="study_ID")
merged_dob_sex_ethnicity_meds  <- merge (load_meds_d_removed, merged_dob_sex_ethnicity, by="study_ID")

merged_dob_sex_ethnicity_meds_no_dup <- merged_dob_sex_ethnicity_meds[duplicated(merged_dob_sex_ethnicity_meds$study_ID)==F,] #Remove duplicates


str(merged_dob_sex_ethnicity_meds_no_dup)
merged_dob_sex_ethnicity_meds$study_ID
merged_dob_sex_ethnicity_meds_no_dup$study_ID
merged_dob_sex_ethnicity_meds[16:22,]

library("Hmisc")
med_describe <- describe(merged_dob_sex_ethnicity_meds_no_dup)
med_describe[c("study_ID","SEX","PHENOTYPE", "Groups_orginal")]


med_describe <- describe(merged_age_sex)
med_describe[c("study_ID","SEX","PHENOTYPE", "Groups_orginal")]
med_describe <- describe(merged_age_sex_ethnicity_meds_duplicate_pass)
med_describe[c("study_ID","SEX","PHENOTYPE", "Groups_orginal")]



dim(merged_age_sex_ethnicity_meds [merged_age_sex_ethnicity_meds$PHENOTYPE =="CASE",])

length(merged_age_sex_ethnicity_meds$SEX)

table(merged_age_sex_ethnicity_meds$SEX)
```



##Add merged dataframe to Lumibatch file. Subset by sampleID first.
```{r}

neweset2 <- neweset[,merged_dob_sex_ethnicity_meds_no_dup$sampleID]
neweset2 <- neweset[,merged_age_sex_ethnicity$sampleID]
str(neweset2)



pData(neweset2) <- merged_dob_sex_ethnicity_meds_no_dup
pData(neweset2) <- merged_age_sex_ethnicity

controlData <- controlData(neweset2)
str(controlData)
controlData$ProbeID
str(neweset2)

## Extra safety check. Not neccesary for analysis, they are just here for you to check what you are doing. 
old_pdata <- pData(neweset2)
head(old_pdata)
head(merged_pheno_african_age_data)

summary(rownames(old_pdata) %in% rownames(merged_pheno_african_age_data) )

dd <- cbind(rownames(old_pdata),rownames(merged_pheno_african_age_data) )
dd <- as.data.frame(dd)
head(dd)
dd$GOOD <- ifelse(dd[,1]==dd[,2], "MATCH","XXX")
table(dd$GOOD)


```

##Save neweset2 to disk
```{r}
?save
save(neweset2, file = paste("~/Documents/A_Year_1_PhD/GAP_Data/Case_Control_Analysis/2014-11-11_case_control_model_Final", "GAP.newset_meds.RData", sep = ""), compress = T)

load ("GAP.neweset_meds.RData")
str (neweset2)
```

## create design matrix
```{r}
gap <- factor(neweset2$PHENOTYPE)
gap_SEX <- factor(neweset2$SEX)
gap_dob <- factor(neweset2$dob)
gap_ethnicitys <- factor (neweset2$ethnicity_selfreportfolders)
gap_african <- as.numeric(neweset2$african)
gap_Indian <- as.numeric(neweset2$Indian)
gap_Age <- as.numeric(neweset2$Age_enhanced)
gap_med_yn <- as.factor (neweset2$yes_no)
gap_med <- as.numeric (neweset2$total_cp)


design = model.matrix(~0+gap+gap_SEX+gap_dob+gap_ethnicitys+gap_med, data=neweset2)

design = model.matrix(~0+gap+gap_SEX+gap_african+gap_Indian+gap_Age+gap_med_yn, data=neweset2)
design = model.matrix(~0+gap+gap_SEX+gap_african+gap_Indian+gap_Age+gap_med, data=neweset2)
design = model.matrix(~0+gap+gap_SEX+gap_african+gap_Indian+gap_Age, data=neweset2)

head(design)
summary (design)



```


## Limma lmFit
```{r limma_lmFit, tidy=TRUE}

fit <- lmFit(neweset2, design)


```

## Limma makeContrasts 

```{r limma_makeContrasts, tidy=TRUE}

contrasts <- makeContrasts(gapCASE-gapCONTROL, levels=design)
contrasts


```



## Limma eBayes on makeContrasts

```{r limma_eBayes_on_makeContrasts, tidy=TRUE}
contrast.fit1 <- contrasts.fit(fit, contrasts)


contrast.fit1 <- eBayes(contrast.fit1)


```

```{r limma_topTable, comment=NA, width=100, tidy=TRUE}
# Limma topTabl
top_de_genes1 <- topTable(contrast.fit1, coef=1, number=5000,adjust.method="fdr",p.value=0.05,confint=TRUE)
head(top_de_genes1)
str(top_de_genes1)


# tidy up
top_de_genes1$logFC <- round(top_de_genes1$logFC,2)
top_de_genes1$CI.025 <- round(top_de_genes1$CI.025,2)
top_de_genes1$CI.975 <- round(top_de_genes1$CI.975,2)
top_de_genes1$AveExpr <- round(top_de_genes1$AveExpr,2)
top_de_genes1$t <- round(top_de_genes1$t,2)
top_de_genes1$P.Value <- signif(top_de_genes1$P.Value,3)
top_de_genes1$adj.P.Val <- signif(top_de_genes1$adj.P.Val,3)
top_de_genes1$B <- round(top_de_genes1$B,2)




# Limma topTable reduce cols
top_de_genes_small1 <- top_de_genes1[,c("SYMBOL","logFC","P.Value","adj.P.Val","B")]

dim(top_de_genes_small1)
head(top_de_genes_small1,  n= 100)
new<-top_de_genes_small1[top_de_genes_small1=="DEFA1B"]



?head
```


## Limma top 100 genes
```{r gene_list}
top_de_genes1$SYMBOL
write.table(top_de_genes1, file="No_Med.txt",row.names=F,quote=FALSE)
write.table(as.data.frame(top_de_genes1$SYMBOL), file="Full_list_Symbol.txt",row.names=F,quote=FALSE,col.names=FALSE)
write.csv(top_de_genes1, file = "all_data_med_ce.csv")

```


```{r gene_list}
Model <- glm( gap ~ gap_SEX+gap_african+gap_Indian+gap_rin+gap_Age+gap_med , family="binomial")
summary(Model)
anova(Model)
```

##Plots
```{r}

x <- subset(top_de_genes_small1,top_de_genes_small1$logFC<0);dim(x) ##for upregulated genes

DEFA1B <-subset (top_de_genes1,top_de_genes1$TargetID=="DEFA1B")
DEFA1B




top_gene_DEFA1B_ISO1 <- "Zp51_cyCF6hIoee3lc"
top_gene_DEFA1B_ISO2 <- "lX5zIIXqEih57eV3es"
top_gene_DEFA1B_ISO3 <- "lV5KVKeJR1KQieOK_c"

gx_gene_DEFA1B_ISO1 <- exprs(neweset2[top_gene_DEFA1B_ISO1,])
gx_gene_DEFA1B_ISO2 <- exprs(neweset2[top_gene_DEFA1B_ISO2,])
gx_gene_DEFA1B_ISO3 <- exprs(neweset2[top_gene_DEFA1B_ISO3,])


plot(as.numeric(gx_gene_DEFA1B_ISO1),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot of DEFA1B_ISO1" )
plot(as.numeric(gx_gene_DEFA1B_ISO2),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot of DEFA1B_ISO2" )
plot(as.numeric(gx_gene_DEFA1B_ISO3),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot of DEFA1B_ISO3" )


par(mfcol=c(2,3))
plot(as.numeric(gx_gene_DEFA1B_ISO1),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot DEFA1B_ISO1", ylab="Relative Expression Level" )
boxplot( as.numeric(gx_gene_DEFA1B_ISO1) ~ neweset2$PHENOTYPE, notch=T, main="Boxplot of DEFA1B_ISO1 ", ylab="Relative Expression Level" )

plot(as.numeric(gx_gene_DEFA1B_ISO2),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot DEFA1B_ISO2", ylab="" )
boxplot( as.numeric(gx_gene_DEFA1B_ISO2) ~ neweset2$PHENOTYPE, notch=T, main="Boxplot of DEFA1B_ISO2 " )

plot(as.numeric(gx_gene_DEFA1B_ISO3),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot DEFA1B_ISO3", ylab="" )
boxplot( as.numeric(gx_gene_DEFA1B_ISO3) ~ neweset2$PHENOTYPE, notch=T, main="Boxplot of DEFA1B_ISO3 " )


##DEFA1 and DEFA3
DEFA1 <-subset (top_de_genes1,top_de_genes1$TargetID=="DEFA1")
DEFA3 <-subset (top_de_genes1,top_de_genes1$TargetID=="DEFA3")
DEFA1
DEFA3

top_gene_DEFA1 <- "QyCF6hIoee3ld3rRXk"
top_gene_DEFA3 <- "cfnMgheoSKHnt5Xd60"

gx_gene_DEFA1 <- exprs(neweset2[top_gene_DEFA1,])
gx_gene_DEFA3 <- exprs(neweset2[top_gene_DEFA3,])

par(mfcol=c(2,2))
plot(as.numeric(gx_gene_DEFA1),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot DEFA1", ylab="Relative Expression Level" )
boxplot( as.numeric(gx_gene_DEFA1) ~ neweset2$PHENOTYPE, notch=T, main="Boxplot of DEFA1 ", ylab="Relative Expression Level" )

plot(as.numeric(gx_gene_DEFA3),col=  ifelse(neweset2$PHENOTYPE=="CASE","red","blue"),main="Scatterplot DEFA3", ylab="" )
boxplot( as.numeric(gx_gene_DEFA3) ~ neweset2$PHENOTYPE, notch=T, main="Boxplot of DEFA3 " )


?boxplot

str(neweset2)

##invalid class "LumiBatch" object: 1: sampleNames differ between assayData and phenoData
##invalid class "LumiBatch" object: 2: sampleNames differ between phenoData and protocolData


```



##Look for DEFA4
```{r}
str(neweset2)

feature_data <- featureData(eset_final)
str(feature_data)
Target_ID<-feature_data$TargetID
Target_ID

find <-Target_ID [Target_ID == "DEFA1B"]

find

write.table(Target_ID, file="ALL_GAP_GENE_SYMBOLS_4756.txt",row.names=F,quote=FALSE)

```


##Check for up and downregulated Genes from Sanders Et all and Gardiner et al. 
```{r}

load_GS_gene_lists <- read.csv("Combined_Gene_list_all_papers.csv", head=TRUE, sep=",",na.string=c("NA","Unknown"),as.is=T)



dim(load_GS_gene_lists[load_GS_gene_lists$ALL])
str(load_GS_gene_lists)
Gap_no_dup_all<- load_GS_gene_lists$GAP_no_DUP_ALL 
Gap_no_dup_all_200<- load_GS_gene_lists$GAP_no_DUP_Top200 
Gap_dup_control<- load_GS_gene_lists$GAP_DUP_CONTROL   
Gap_dup_control_200<- load_GS_gene_lists$GAP_DUP_Control_Top200
Gap_dup<- load_GS_gene_lists$GAP_DUP
all_gap <- top_de_genes1$SYMBOL


gardiner_data <- load_GS_gene_lists$gardiner
sanders_data <- load_GS_gene_lists$sanders
all_data <- load_GS_gene_lists$All

length(all_gap[all_gap %in% all_data])
all_gap[all_gap %in% all_data]

length(Gap_dup_control[Gap_dup_control %in% all_data])
Gap_dup_control[Gap_dup_control %in% all_data]

length(Gap_dup[Gap_dup %in% all_data])
Gap_dup[Gap_dup %in% all_data]
Gap_dup

Gap_dup_control[Gap_dup_control%in%all_data]
Gap_dup[Gap_dup%in%all_data]

```

##Volcano Plot
```{r}

##ggplot2 volcanoplot
library (ggplot2)
top_de_genes_vol <- topTable(contrast.fit1, coef=1, number=5000,adjust.method="fdr",p.value=1,confint=TRUE, sort.by="logFC")
top_de_genes_vol$threshold = as.factor(abs(top_de_genes_vol$logFC) > 0.0 & top_de_genes_vol$adj.P.Val < 0.05)
str(top_de_genes_vol)


G_volcanoplot = ggplot(data=top_de_genes_vol, aes(x=logFC, y=-log10(P.Value), colour=threshold)) +
  geom_point(alpha=0.4, size=1.75) +
  theme(legend.position = "none") +
  xlim(c(-1.5, 1.5)) + ylim(c(0, 7.5)) +
  xlab("Relative change in Gene expression") + ylab("-log10 unadjusted p-value")+
  geom_vline(xintercept = 0, colour = "black")+
  ggtitle ("Volcano Plot of GAP expression Data")
  ##theme(plot.title = element_text(lineheight "12", face="bold"))
G_volcanoplot 


# Installation ------------------------------------------------------------
 
# First, install devtools from CRAN then install Tmisc 
# directly from GitHub using devtools (https://github.com/stephenturner/Tmisc)
install.packages("calibrate")
devtools::install_github("stephenturner/Tmisc")
 
# Load the Tmisc library

str(top_de_genes_vol)

# Make a basic volcano plot
library(Tmisc)
library(calibrate)

with(top_de_genes_vol, plot(logFC, -log10(P.Value), pch=20, main="Volcano plot (12_ethnicity_med_model)", xlim=c(-1.5,1.5)))

# Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(top_de_genes_vol, adj.P.Val >0.05 ), points(logFC, -log10(P.Value), pch=20, col="firebrick"))
with(subset(top_de_genes_vol, adj.P.Val <0.05 ), points(logFC, -log10(P.Value), pch=20, col="navy"))
##with(subset(top_de_genes_vol, logFC >1 ), points(logFC, -log10(P.Value), pch=20, col="green"))

 
# Label points with the textxy function from the calibrate plot
with(subset(top_de_genes_vol, logFC >0.65), textxy(logFC, -log10(P.Value), labs=TargetID, cex=.7))
with(subset(top_de_genes_vol, logFC < -0.6& adj.P.Val <0.05 ), textxy(logFC, -log10(P.Value), labs=TargetID, cex=.7))
with(subset(top_de_genes_vol, adj.P.Val <0.0006), textxy(logFC, -log10(P.Value), labs=TargetID, cex=.7))
#with(subset(top_de_genes_vol, logFC < -0.6 & adj.P.Val <0.008 ), textxy(logFC, -log10(P.Value), labs=TargetID, cex=.7))


```