---
title: "GAP_GX_rlm"
author: "Daniel Leirer"
date: "07/01/2015"
output: html_document
---

PCA Batch Regressions
======================
making this a requirement ie if [tech_pheno_file] exits then test for assoc of with PC1 and batches etc


## 0. Load Relevant Data


## 1. Set up data for batch regressions
```{r start_PCA_Batch_Regressions, eval=TRUE}
# a little fix/renaming
eset_lumiN <- eset_bg_log2_rsn 
cat(" Starting batch versus PC1 and PHENOTYPE Rregressions [PC1 ~ batch_var]","\r")
cat(" Getting Gene expression matrix ","\r")
gx <- exprs(eset_lumiN)
gx <- t(gx)
cat(" Reading in technical information on eg [Sample.ID, RIN, RNA_YIELD, BATCH, CHIP, DATE_CHIP_RUN, DATE_RNA_EXTRACTED] ",tech_pheno_file ,"\r")
tech_pheno <- read.table(paste(tech_pheno_file),head=TRUE,sep=" ") ## Sample.ID,RIN,RNA_YIELD,BATCH,CHIP, DATE_CHIP_RUN,DATE_RNA_EXTRACTED
str(tech_pheno)
tech_pheno$Sentrix.Barcode <- as.factor(tech_pheno$Sentrix.Barcode)
pdata <- pData(eset_lumiN)
pdata <- as.data.frame(pdata[,c("sampleID","Index")])
colnames(pdata) <- c("sampleID","Index")
tech_batch <- merge( pdata, tech_pheno, by.x="sampleID", by.y="Sample.ID", sort=FALSE, all.x=TRUE)
tech_batch <- tech_batch[order(tech_batch$Index),]
tech_batch <- tech_batch[,3:dim(tech_batch)[2]]
head(tech_batch)
str(tech_batch)
# get names of var
cat(" get names of var ","\r")
batch_var_names <- names(tech_batch)
date_vars <- grep("Date",batch_var_names) # which ones are dates
batch_var_names
## Run PCA
cat(" Running PCA on t(exprs(eset)) ","\r")
pca_gx <- prcomp(gx)$x
pca_gx <- pca_gx[,"PC1"]
# PHENOTYPES FOR REGRESSIONS
cat(" setting up phenotypes PC1,PHENOTYPE & GROUPS for regressions ","\r")
PC1 <- as.numeric(pca_gx)
PHENOTYPE <- as.numeric(as.factor(toupper(pData(eset_lumiN)$PHENOTYPE))) # toupper() called because of pesky 'case' issues
GROUPS <-    as.numeric(as.factor(toupper(pData(eset_lumiN)$GROUPS))) # toupper() called because of pesky 'case' issues
#
#df <- cbind(tech_batch,PC1,PHENOTYPE,GROUPS)
#df_z <- apply(df,2,as.factor)
#df_z <- apply(df_z,2,as.numeric)
str(pca_gx)


```

## 2. multivariate model
```{r multivariate_model_PCA_Batch_Regressions, eval=TRUE}
# Test for association of batch vars with PC1
# multivariate full model
multivariate_model_terms <- as.formula(paste("PC1 ~ ", paste(batch_var_names ,collapse="+")))
#########################
# PC1 is run last 


## Doing rlm

## Step 0 : Load Roboustbase
library(robustbase)

## Step 1 : fit the full model

# Drop empty factor levels by saving and loading.

str(tech_batch)
table(tech_batch$Sentrix.Barcode)

write.table(tech_batch, file=paste(out_dir,"/",project_name,".tech_batch_factor_clear.txt",sep=""),sep="\t")
tech_batch4 <- read.table(file=paste(out_dir,"/",project_name,".tech_batch_factor_clear.txt",sep=""), sep="\t")
str(tech_batch4)

tech_batch4$Sentrix.Barcode <- as.factor(tech_batch4$Sentrix.Barcode)
tech_batch4$SampleSection <- as.factor(tech_batch4$SampleSection)

robust_m1  <- lmrob(PC1 ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))
robust_m1  <- lmrob(PC1 ~ Sentrix.Barcode, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

linear_m1  <- lm(PC1 ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4)

## Step 2: See whats significant at t< +/- 1.96


summary(robust_m1)
summary(linear_m1)
anova(linear_m1)

## Step 3 : Remove non sig terms and re-fit the model (using update() funnction)
str(tech_batch4)

robust_m2  <- lmrob(PC1 ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + labelled_cRNA_Yield, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant term (1.43e-07 and 1.43e-07) 


summary(robust_m2)



## repeat until all terms are significant
## final list of covar is what we use in batch regressions
## the AIM is to find which batch variables are associated with PC1. 


```

Batch Correction using linear models
=====================================
```{r linear_model_batch_correction_prep, eval=TRUE}



####
if(pheno!="PC1") stop(" WARNING!: model terms are not from the PC1 ~ batch regressions" )

# get gene expressuion
gx <- exprs(eset_bg_log2_rsn)
n_probes <- dim(gx)[1]
gx <- t(gx)
dim(gx)
gx[1:10,1:10]

# probe nuID names
new_probe_names <- paste("p_",colnames(gx),sep="") # ADD p as some nuID start with a number
##test_new_probe_names <- new_probe_names [1:5]
head(new_probe_names)
#probe_names <- colnames(gx)
head(new_probe_names)
colnames(gx) <- new_probe_names
gx[1:10,1:10]

# make new matrix to write adjusted values to
adj_gx <- gx*0
adj_gx[1:10,1:10]
##test_adj_gx<- adj_gx[1:10,1:5]
##test_adj_gx
# get batch phenos
##batch_pheno <- tech_batch[,anova_data$terms] ##Can't create batch_pheno do not have anova_data. anova_data$terms is simply the significant variables

##batch_pheno2 <- tech_batch4[,c("Sentrix.Barcode", "Date_out", "Conc_Nanodrop", "Date_extraction", "labelled_cRNA_Yield")]
##batch_pheno2 <- cbind(batch_pheno2,gx)
# this is the data for the regression
batch_pheno2[1:10,1:10]
##str(batch_pheno2)
#
eset_bg_log2_rsn_regression_input <- batch_pheno2

save(eset_bg_log2_rsn_regression_input,file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.regression_input.RData",sep=""))

##test_batch_pheno2 <- batch_pheno2[1:10,1:10]

##write.table(test_batch_pheno2,file=paste(out_dir,"/",project_name,".test_batch_pheno2.txt",sep=""))
##test_batch_pheno2 <- read.table("GAP_DL_FINAL.test_batch_pheno2.txt",head=TRUE,sep="")

##str(test_batch_pheno2)
##test_batch_pheno2
##table(test_batch_pheno2$Sentrix.Barcode)


```

## THIS IS THE ALTERNATIVE TO LINEAR AND ROBUST LINEAR REGGRESSIONS USING COMBAT (use tech_batch4)
```{r COMBAT, eval=TRUE}
####################### START HERE #################################


## Step 0 : Load Roboustbase, ComBat and tech_batch4. (also make sure all relevant covariates are factors)
library(robustbase)
library(sva)



## Step 1 : Adjust for Sentrix.Barcode

#### Step 1.1 : define batch, model matrix and invert expression data matrix 

gx_t <- t(gx)

gx [1:10,1:10]
gx_t [1:10,1:10]

batchSB <- tech_batch4$Sentrix.Barcode
modcombat3 <- model.matrix(~1, data=tech_batch4) 


#### Step 1.2 : Run ComBat

combat_edata_2 <- ComBat(dat=gx_t, batch=batchSB, mod=modcombat3, numCovs=NULL, par.prior=TRUE, prior.plots=FALSE)

combat_edata_2 [1:10,1:10]

#### Step 1.3 : Generate Princple Components of ComBat adjusted expression matrix
combat_edata_2 <- t(combat_edata_2)
pca_combat_edata_2 <- prcomp(combat_edata_2)$x
pca_combat_edata_2 <- pca_combat_edata_2[,"PC1"]
str(pca_combat_edata_2)
str(pca_gx)

PC1_SB <- as.numeric(pca_combat_edata_2)


#### Step 1.4 : Check is Sentrix Barcode is still associated with the new PC1. Then check if anything in the full model is associated. Then check if anything in the reduced model only containing the previously significant covariates is still significant. 

robust_m1_SB  <- lmrob(PC1_SB ~ Sentrix.Barcode, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_SB) ## No significant

robust_m1_SB  <- lmrob(PC1_SB ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_SB) ## The following are significant here : Sentrix.Barcode, Date_out, Conc_nanodrop, Date_extraction
str(tech_batch4)

robust_m1_SB  <- lmrob(PC1_SB ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + concentration_of_labelled_cRNA, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

summary(robust_m1_SB) ## All

robust_m1_SB  <- lm(PC1_SB ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + concentration_of_labelled_cRNA, data=tech_batch4) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

anova(robust_m1_SB)


######################################### Step 2 ##########################################

## Step 2 : Adjust for Date_out

#### Step 2.1 : define batch, model matrix and invert expression data matrix 

combat_edata_2_t <- t(combat_edata_2)

combat_edata_2 [1:10,1:10]
combat_edata_2_t [1:10,1:10]

batchDO <- tech_batch4$Date_out
modcombat <- model.matrix(~1, data=tech_batch4) 
str(batchDO)

#### Step 2.2 : Run ComBat

combat_edata_DO <- ComBat(dat=combat_edata_2_t, batch=batchDO, mod=modcombat, numCovs=NULL, par.prior=TRUE, prior.plots=FALSE)

combat_edata_DO [1:10,1:10]

#### Step 2.3 : Generate Princple Components of ComBat adjusted expression matrix
combat_edata_DO <- t(combat_edata_DO)
pca_combat_edata_DO <- prcomp(combat_edata_DO)$x
pca_combat_edata_DO <- pca_combat_edata_DO[,"PC1"]
str(pca_combat_edata_DO)
str(pca_combat_edata_2)
str(pca_gx)

PC1_DO <- as.numeric(pca_combat_edata_DO)


#### Step 2.4 : Check is Sentrix Barcode is still associated with the new PC1. Then check if anything in the full model is associated. Then check if anything in the reduced model only containing the previously significant covariates is still significant. 

robust_m1_DO  <- lmrob(PC1_DO ~ Date_out, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_DO) ## No significant

robust_m1_DO  <- lmrob(PC1_DO ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_DO) ## The following are significant here : Sentrix.Barcode, Conc_nanodrop, Date_extraction
str(tech_batch4)

robust_m1_DO  <- lmrob(PC1_DO ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + labelled_cRNA_Yield, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

robust_m1_DO  <- lm(PC1_DO ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + labelled_cRNA_Yield, data=tech_batch4)

summary(robust_m1_DO) ## The following are significant here : Intercept, Date_out, Conc_nanodrop, Date_extraction, concentration_of_labelled_cRNA, labelled_cRNA_Yield
anova(robust_m1_DO)

######################################### Step 3 ##########################################

## Step 3 : Adjust for Date_extraction

#### Step 3.1 : define batch, model matrix and invert expression data matrix 

combat_edata_DO_t <- t(combat_edata_DO)

combat_edata_DO [1:10,1:10]
combat_edata_DO_t [1:10,1:10]

batchDE <- tech_batch4$Date_extraction
modcombat <- model.matrix(~1, data=tech_batch4) 
str(batchDE)

#### Step 3.2 : Run ComBat

combat_edata_DE <- ComBat(dat=combat_edata_DO_t, batch=batchDE, mod=modcombat, numCovs=NULL, par.prior=TRUE, prior.plots=FALSE)

combat_edata_DE [1:10,1:10]

#### Step 3.3 : Generate Princple Components of ComBat adjusted expression matrix
combat_edata_DE <- t(combat_edata_DE)
pca_combat_edata_DE <- prcomp(combat_edata_DE)$x
pca_combat_edata_DE <- pca_combat_edata_DE[,"PC1"]
str(pca_combat_edata_DE)
str(pca_combat_edata_DO)
str(pca_combat_edata_2)
str(pca_gx)

PC1_DE <- as.numeric(pca_combat_edata_DE)


#### Step 3.4 : Check is Sentrix Barcode is still associated with the new PC1. Then check if anything in the full model is associated. Then check if anything in the reduced model only containing the previously significant covariates is still significant. 

robust_m1_DE  <- lmrob(PC1_DE ~ Date_extraction, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_DE) ## No significant

robust_m1_DE  <- lmrob(PC1_DE ~ Sentrix.Barcode + Date_out + Date_extraction, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_DE) ## No significant



robust_m1_DE  <- lmrob(PC1_DE ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_DE) ## The following are significant here : Sentrix.Barcode, Conc_nanodrop


robust_m1_DE  <- lmrob(PC1_DE ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + concentration_of_labelled_cRNA, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

robust_m1_DE  <- lmrob(PC1_DE ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + concentration_of_labelled_cRNA, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

robust_m1_DE  <- lmrob(PC1_DE ~ Conc_Nanodrop + concentration_of_labelled_cRNA, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

cor.test (tech_batch4$Conc_Nanodrop, tech_batch4$concentration_of_labelled_cRNA)
plot(tech_batch4$Conc_Nanodrop, tech_batch4$concentration_of_labelled_cRNA)

summary(robust_m1_DE) ## The following are significant here : Intercept, Date_out, Conc_nanodrop, concentration_of_labelled_cRNA, labelled_cRNA_Yield

linear_m1_DE  <- lm(PC1_DE ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4,)

summary(linear_m1_DE) ## The following are significant here : Sentrix.Barcode, Conc_nanodrop
anova(linear_m1_DE)


```



## SET UP FOR ROBOUST LINEAR REGRESSION
```{r linear_model_batch_correction_prep, eval=TRUE}



####
if(pheno!="PC1") stop(" WARNING!: model terms are not from the PC1 ~ batch regressions" )

# get gene expressuion
gx <- exprs(eset_bg_log2_rsn)
n_probes <- dim(gx)[1]
gx <- t(gx)
dim(gx)
gx[1:10,1:10]

combat_edata_DE_t <- t(combat_edata_DE) ## From Sentrix.Barcode, Date_out and Date_extraction ComBat corrected 
combat_edata_DE_t [1:10,1:10]

# probe nuID names
##new_probe_names <- paste("p_",colnames(combat_edata_DE_t),sep="") # ADD p as some nuID start with a number

#probe_names <- colnames(gx)
head(new_probe_names)
colnames(combat_edata_DE) <- new_probe_names
combat_edata_DE[1:10,1:10]

# make new matrix to write adjusted values to
adj_gx <- combat_edata_DE*0
adj_gx[1:10,1:10]


batch_pheno_C <- tech_batch4[,c( "Conc_Nanodrop", "concentration_of_labelled_cRNA")]
batch_pheno_C <- cbind(batch_pheno_C,combat_edata_DE)
# this is the data for the regression
batch_pheno_C[1:10,1:10]
str(batch_pheno_C)
#
eset_bg_log2_rsn_regression_input_combat_adj <- batch_pheno_C

save(eset_bg_log2_rsn_regression_input_combat_adj,file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.regression_input_combat_adj.RData",sep=""))


```

## THIS ROBOUST LINEAR REGRESSION FOR THE FULL DATA AFTER SIGNIFICANT FACTORIAL BATCHES HAVE BEEN ELIMINATED. IT WILL LIEKLY TAKE HOURS TO PROCESS (approx 5.5 hours)
## (covariates left are Conc_Nanodrop and either labelled_cRNA_Yield or concentration_of_labelled_cRNA. concentration_of_labelled_cRNA will be used)

```{r linear_model_batch_correction_run, eval=TRUE}
# loop through each probe and adjust for sig batches
pn <- 1
#
for(probe in new_probe_names){
    

    rlm_model <- as.formula( paste(probe,"~","Conc_Nanodrop + concentration_of_labelled_cRNA",sep="") )
    rlm_probe  <- lmrob(rlm_model, data=batch_pheno_C, control = lmrob.control(setting = "KS2014")) ##Probe instead of PC1 right??? summary(rlm_probe)
  
    rsq <- round( summary(rlm_probe)$adj.r.squared, 3)   
    residual_probe <- rlm_probe$residual 
    
    mean_probe_level <- mean(batch_pheno_C[,probe])
    
    adjusted_probe_level <- residual_probe + mean_probe_level
    
    adj_gx[,probe] <- adjusted_probe_level
    
   ### cat("  Progress: ",pn," : ",round(pn/n_probes,3),"\r")
    
    sink(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.rlm_probe_progress_rsq_combat_adj.txt",sep=""),append=TRUE) ##changed lm to rlm
    
    cat(" doing [",probe ,"] ~ [","Conc_Nanodrop + concentration_of_labelled_cRNA","].RSQ=",rsq,". Progress:",round(pn/n_probes,3),"\r","\n")
    
    sink()
        
    pn <- pn + 1

    }




# update names and transform back to probe x sample matrix
adj_gx <- t(adj_gx)
rownames(adj_gx) <- rownames(exprs(eset_bg_log2_rsn))
adj_gx[1:10,1:10]

# save raw matrix
eset_bg_log2_rsn_adj_gx_combat_adj <- adj_gx
save(eset_bg_log2_rsn_adj_gx_combat_adj,file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj_gx_combat_adj.RData",sep=""))

```



##  CHECK IF eset_bg_log2_rsn_adj_gx_combat_adj STILL HAS SIGNIFICANT ASSOCIATIONS 
```{r, eval=TRUE}

#### Step 1 : Generate Princple Components of ComBat adjusted expression matrix
eset_bg_log2_rsn_adj_gx_combat_adj_t <- t(eset_bg_log2_rsn_adj_gx_combat_adj)
eset_bg_log2_rsn_adj_gx_combat_adj_t <- prcomp(eset_bg_log2_rsn_adj_gx_combat_adj_t)$x
pca_eset_bg_log2_rsn_adj_gx_combat_adj_t <- eset_bg_log2_rsn_adj_gx_combat_adj_t[,"PC1"]
str(pca_eset_bg_log2_rsn_adj_gx_combat_adj_t)
str(pca_combat_edata_DE)
str(pca_combat_edata_DO)
str(pca_combat_edata_2)
str(pca_gx)

PC1_CAF <- as.numeric(pca_eset_bg_log2_rsn_adj_gx_combat_adj_t) ##COMBAT ADJUSTED FINAL


#### Step 2 : Check if Conc_Nanodrop + concentration_of_labelled_cRNA are still associated with the new PC1. Then check if anything in the full model is associated. Then check if anything in the reduced model only containing the previously significant covariates is still significant. 





robust_m1_CAF  <- lmrob(PC1_CAF ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))



robust_m1_CAF  <- lmrob(PC1_CAF ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4, control = lmrob.control(setting = "KS2014"))

summary(robust_m1_CAF) ## The following are significant here : Sentrix.Barcode


robust_m1_CAF  <- lmrob(PC1_CAF ~ Sentrix.Barcode + Date_out + Conc_Nanodrop + Date_extraction + concentration_of_labelled_cRNA, data=tech_batch4, control = lmrob.control(setting = "KS2014")) ## both labelled_cRNA_Yield and concentration_of_labelled_cRNA are significant terms (1.43e-07 and 1.43e-07). Probably interchangable 

summary(robust_m1_CAF) ## The following are significant here : Date_out, concentration_of_labelled_cRNA



linear_m1_CAF  <- lm(PC1_CAF ~ RIN + Sentrix.Barcode + SampleSection + Date_out + Batch  + person + Conc_Nanodrop + labelled_cRNA_Yield + concentration_of_labelled_cRNA + Date_extraction + Date_Quantitation_by_RiboGreen + Date_labelled_cRNA + Date_cRNApurification + Date_Dilutionand_Amplification, data=tech_batch4,)

summary(linear_m1_CAF) ## The following are significant here : Sentrix.Barcode
anova(linear_m1_CAF) ## The following are significant here Batch, labelled_cRNA_Yield


```


## CHECK eset_bg_log2_rsn_adj_gx_combat_adj HAS COMPLETE MATRIX. ALSO CHECK BY SVA IF ANY HIDDEN VARIABLES
```{r , eval=TRUE}

load("GAP_DL_FINAL.eset_bg_log2_rsn_adj_gx_combat_adj.RData")

dim(eset_bg_log2_rsn_adj_gx_combat_adj)
eset_bg_log2_rsn_adj_gx_combat_adj [47220:47231,480:487]

## SVA

eset_lumiN <- eset_bg_log2_rsn 
gx <- exprs(eset_lumiN)
gx <- t(gx)
gx_t = t(gx) 

pheno_data <- pData(eset_bg_log2_rsn_adj)
str(pheno_data)

edata_o =  gx_t
edata = eset_bg_log2_rsn_adj_gx_combat_adj


edata [1:10,1:10]

mod = model.matrix(~as.factor(GROUPS), data=pheno_data)

n.sv_o = num.sv(edata_o,mod,method="be")
n.sv_o ## Using be 39 using leek 0
n.sv = num.sv(edata,mod,method="be")
n.sv ## Using be 54 Using leek 0

```

Make Batch Adjusted Data Set
=================================================

```{r make_batch_adjusted_data, eval=TRUE}
# make new eset and replace exprs() matrix with new batch adjusted data
eset_bg_log2_rsn_adj <- eset_bg_log2_rsn
exprs(eset_bg_log2_rsn_adj) <- eset_bg_log2_rsn_adj_gx_combat_adj
str(eset_bg_log2_rsn_adj)

```

```{r save_batch_adjusted_data, eval=TRUE}
# save eset_bg_log2_rsn_adj
save(eset_bg_log2_rsn_adj, file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj.RData",sep="")  , compress=T)
```

```{r write_expression_files_batch_adjusted_data, eval=TRUE}
# write_expression_files eset_bg_log2_rsn_adj
write_expression_files(eset=eset_bg_log2_rsn_adj,outfile= paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj",sep="") )
```


QC Plots of `eset_bg_log2_rsn_adj`
---------------------------------------------------------------
## basic_qc_plot_lumi 
```{r basic_qc_plot_lumi_eset_bg_log2_rsn_adj, fig.height=7, fig.width=10,eval=TRUE}
# basic plots
# plot to screen


basic_qc_plot_lumi(eset_bg_log2_rsn_adj)
par(def.par)
```

```{r basic_qc_plot_lumi_eset_bg_log2_rsn_adj_pdf, echo=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj.basic_qc_plot_lumi.pdf",sep=""),width=11,height=8)
basic_qc_plot_lumi(eset_bg_log2_rsn_adj)
dev.off()
par(def.par)

```

## coloured_dendrogram_lumi 
```{r coloured_dendrogram_lumi_eset_bg_log2_rsn_adj, fig.height=7, fig.width=10,eval=TRUE}
# coloured_dendrogram_lumi
# plot to screen
par(c=5,20,5,5)
coloured_dendrogram_lumi(eset_bg_log2_rsn_adj)
par(def.par)
```

```{r coloured_dendrogram_lumi_eset_bg_log2_rsn_adj_pdf,eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj.coloured_dendrogram_lumi.pdf",sep=""),width=11,height=8)
coloured_dendrogram_lumi(eset_bg_log2_rsn_adj)
dev.off()
par(def.par)

```

## pca_plot_lumi 
```{r pca_plot_lumi_eset_bg_log2_rsn_adj, fig.height=7, fig.width=7,eval=TRUE}
# PCA plots
# plot to screen
pca_plot_lumi(eset_bg_log2_rsn_adj)
par(def.par)
```

```{r pca_plot_lumi_eset_bg_log2_rsn_adj_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj.pca_plot_lumi.pdf",sep=""),width=7,height=7)
pca_plot_lumi(eset_bg_log2_rsn_adj)
dev.off()
```
## SampleNetwork Plots eset_bg_log2_rsn_adj
```{r sampleNetwork_plot_all_lumi_eset_bg_log2_rsn_adj, fig.height=10, fig.width=10,eval=TRUE}
# SampleNetwork Plots
# plot to screen
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_adj, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_adj, colBy="group");
par(def.par)
```

```{r sampleNetwork_plot_all_lumi_eset_bg_log2_rsn_adj_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_adj.sampleNetwork_plot_all_lumi.pdf",sep=""),width=8,height=8)
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_adj, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_adj, colBy="group");
dev.off()
```


****

Create Final QC'd Expression data set: Good Probes & Good Samples
==================================================================
Subset data to good_probes.   
At this stage we have already removed sample outliers and adjusted for significant batch variables. Now we subset to probes that can be reliably detected in 80% of each GROUP. This is the final data set to be used on all analyses. 
Some of the QC plot wont look as nice, eg CV plots, as these genes should represent some real biology.

```{r create_final_eset, eval=TRUE}
# subset to good probes 


eset_final <- eset_bg_log2_rsn_adj[good_probes,]
str(eset_final)

```

```{r save_final_eset, eval=TRUE}
# subset to good probes 
save(eset_final, file=paste(out_dir,"/",project_name,".eset_final.RData",sep="")  , compress=T)
```

```{r write_expression_files_final_eset, eval=TRUE}
# write_expression_files eset_bg_log2_rsn_adj
write_expression_files(eset=eset_final,outfile= paste(out_dir,"/",project_name,".eset_final",sep="") )
```

QC Plots of `eset_final`
---------------------------------------------------------------
## basic_qc_plot_lumi 
```{r basic_qc_plot_lumi_eset_final, fig.height=7, fig.width=10,eval=TRUE}
# basic plots
# plot to screen
basic_qc_plot_lumi(eset_final)
par(def.par)
```

```{r basic_qc_plot_lumi_eset_final_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_final.basic_qc_plot_lumi.pdf",sep=""),width=11,height=8)
basic_qc_plot_lumi(eset_final)
dev.off()
```

## coloured_dendrogram_lumi 
```{r coloured_dendrogram_lumi_eset_final, fig.height=7, fig.width=10, eval=TRUE}
# coloured_dendrogram_lumi
# plot to screen
par(mar=c(5,20,5,5))
coloured_dendrogram_lumi(eset_final)
par(def.par)
```

```{r coloured_dendrogram_lumi_eset_final_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_final.coloured_dendrogram_lumi.pdf",sep=""),width=11,height=8)
coloured_dendrogram_lumi(eset_final)
dev.off()
```

## pca_plot_lumi 
```{r pca_plot_lumi_eset_final, fig.height=7, fig.width=7, eval=TRUE}
# PCA plots
# plot to screen
pca_plot_lumi(eset_final)
par(def.par)
```

```{r pca_plot_lumi_eset_final_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_final.pca_plot_lumi.pdf",sep=""),width=7,height=7)
pca_plot_lumi(eset_final)
dev.off()
```

## SampleNetwork Plots 
```{r sampleNetwork_plot_all_lumi_eset_final, fig.height=10, fig.width=10,eval=TRUE}
# SampleNetwork Plots
# plot to screen
sampleNetwork_plot_all_lumi(eset_final, colBy="chip");
sampleNetwork_plot_all_lumi(eset_final, colBy="group");
par(def.par)
```

```{r sampleNetwork_plot_all_lumi_eset_final_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_final.sampleNetwork_plot_all_lumi.pdf",sep=""),width=8,height=8)
sampleNetwork_plot_all_lumi(eset_final, colBy="chip");
sampleNetwork_plot_all_lumi(eset_final, colBy="group");
dev.off()
```

****

Clean up
=========

```{r clean_up, eval=TRUE}
cat(" Clean up processing directory","\r","\n")

cat(" Making data directories","\r","\n")
system( paste(" mkdir ",out_dir,"/eset_raw",sep=""));
system( paste(" mkdir ",out_dir,"/eset_bg",sep=""));
system( paste(" mkdir ",out_dir,"/eset_bg_log2_rsn",sep=""));
system( paste(" mkdir ",out_dir,"/eset_final",sep=""));
system( paste(" mkdir ",out_dir,"/XIST_Gender_checks",sep=""));
system( paste(" mkdir ",out_dir,"/detected_probes",sep=""));
system( paste(" mkdir ",out_dir,"/SampleNetwork",sep=""));
system( paste(" mkdir ",out_dir,"/batch_regressions/",sep=""));

cat(" Cleaning up data directories","\r","\n")
# XIST
system( paste(" mv -v ",out_dir,"/","*.XIST* ",out_dir,"/XIST_Gender_checks/",sep=""))

# detected probes
system( paste(" mv -v ",out_dir,"/","*.detected_probes* ",out_dir,"/detected_probes/",sep=""))

# SampleNetwork
system( paste("  mv -v ",out_dir,"/","*ampleNetwork* ",out_dir,"/SampleNetwork/",sep=""))

# batch regressions multivariate
system( paste(" mv -v ",out_dir,"/","*multivariate* ",out_dir,"/batch_regressions/",sep=""))

# eset_****
system( paste(" mv -v ",out_dir,"/","*.eset_raw.* ",out_dir,"/eset_raw/",sep=""))
system( paste(" mv -v ",out_dir,"/","*.eset_bg.* ",out_dir,"/eset_bg/",sep=""))
system( paste(" mv -v ",out_dir,"/","*.eset_bg_log2_rsn* ",out_dir,"/eset_bg_log2_rsn/",sep=""))
system( paste(" mv -v ",out_dir,"/","*.eset_final.* ",out_dir,"/eset_final/",sep=""))

```