****

Id Detected Probes per GROUPS
==============================
Here we are creating lists of probes that have expression levels greater than of the mean intensity of the negative control beads.  
This seems to be a better measure of "expressed/detected" than Illumina's own detection p-values. You can see proof of that when looking at XIST expression levels versus Geneder using Illumina's own detection p-values for both p=0.05 and p=0.01!  
The expression levels are taken from the **background corrected** data \[eset_bg\].
This needs be be run after the gender check, as Y Chrom probe expression is determined in xist_MALES only. 

```{r id_detected_probes_per_group, eval=TRUE}
## PROBE DETECTED 2SD ABOVE MEAN BACKGROUND ##
cat(" Calculating Probe Detection rates. \n Probe is seen as Detected if it has background corrected signal intensity greather than the mean intensity of the negative control beads","\r","\n")
## get expression matrix
## gx <- exprs(eset_bg)
## chaned to raw measures as this brings some more data
gx <- exprs(eset_raw)


## get negative bead ranges mean or max or 2SD mean  of neg beads
neg_2sd <- neg_mean + 2*neg_sd

## sweep through gx matrix to id probes greater than Mean of negative beads
## rows are probes, cols are samples 

## THIS IS THE MAIN PROBE DETECTION CALCULATION
det <- sweep(gx, 1, round(neg_mean,2),">")

## Writing Probe Detection Calls to file #
det_tmp  <- as.data.frame(det)
det_tmp$nuID <- rownames(det_tmp)
det_tmp$min_expression <- apply(gx,1,min)
det_tmp$max_expression <- apply(gx,1,max)
det_tmp$mean_expression <- apply(gx,1,mean)
det_tmp$sd_expression <- apply(gx,1,sd)

probe_detected <- merge(fData(eset_bg), det_tmp , by.x="nuID",by.y="nuID",sort=FALSE)

cat(" Writing Probe Detection Calls to",paste(out_dir,"/",project_name,".eset_bg.probe_detected.txt",sep="") ,"\r","\n")
write.table(probe_detected , file=paste(out_dir,"/",project_name,".eset_bg.probe_detected.txt",sep=""),sep="\t",row.names=FALSE, quote=FALSE)

## probe_detection counts
probe_detection <- rowSums(det)

## n samples
n_samples <- dim(gx)[2]

## probe annotations
probes_not_detected_in_any_sample <- probe_detection==0
probes_detected_in_50_sample <- probe_detection>=n_samples*0.50
probes_detected_in_80_sample <- probe_detection>=n_samples*0.80
probes_detected_in_all_sample <- probe_detection==n_samples
# 
probe_annotations_0_detected <- fData(eset_bg[probes_not_detected_in_any_sample,])
probe_annotations_50_detected  <- fData(eset_bg[probes_detected_in_50_sample ,])
probe_annotations_80_detected  <- fData(eset_bg[probes_detected_in_80_sample ,])
probe_annotations_100_detected  <- fData(eset_bg[probes_detected_in_all_sample,])

cat(" Adding detetion call rate for all probes and samples to fData() slot for eset_bg","\r","\n")
fData(eset_bg)$n_detected <- probe_detection
fData(eset_bg)$n_detected_call_rate <- round( probe_detection/n_samples ,3)
fData(eset_bg)$probes_not_detected_in_any_sample <- probe_detection==0
fData(eset_bg)$probes_detected_in_50_sample <- probe_detection>=n_samples*0.50
fData(eset_bg)$probes_detected_in_80_sample <- probe_detection>=n_samples*0.80
fData(eset_bg)$probes_detected_in_all_sample <- probe_detection==n_samples
# add min, max, mean, sd, median
fData(eset_bg)$min_expression <- round(apply(exprs(eset_bg),1,min),3)
fData(eset_bg)$max_expression <- round(apply(exprs(eset_bg),1,max),3)
fData(eset_bg)$mean_expression <- round(apply(exprs(eset_bg),1,mean),3)
fData(eset_bg)$sd_expression <- round(apply(exprs(eset_bg),1,sd),3)
fData(eset_bg)$median_expression <- round(apply(exprs(eset_bg),1,median),3)

## sample_detection counts
cat(" sample probe detection rate","\r","\n")
n_probes <- dim(eset_bg)[1]
sample_detection <- colSums(det)
pData(eset_bg)$n_probes_detected <- sample_detection
pData(eset_bg)$n_probes_detected_call_rate <- round( sample_detection/n_probes ,3)
save(eset_bg, file=paste(out_dir,"/",project_name,".eset_bg.RData",sep=""))

## get group information from pData() slot
group_names <- unique(pData(eset_bg)$GROUPS);
group_names
groups  <- pData(eset_bg)$GROUPS
n_groups <- length(group_names)

## get expression matrix ##
gx <- exprs(eset_bg)
# get neg_mean values. Calculated previously
head(neg_mean)

#############################################################################################
## THIS IS THE MAIN PROBE DETECTION CALCULATION
## loop through each group and id probes greater than  mean neg beads in X% of samples/group
#############################################################################################

for(n in group_names ) {
cat(" Finding probes in ",probe_det/100," of sample group [",n,"] with signal intensity greather than mean intensity of the negative control beads ","\r","\n")

group_label <- paste(n)

sel_samples <- pData(eset_bg)$GROUPS==n;

n_samples_in_group <- dim(gx[,sel_samples])[2];

cat(" Number of samples in group [",n,"] = ",n_samples_in_group,"\r","\n")

detection_matrix <- sweep(gx[,sel_samples],1,round(neg_mean,2)[sel_samples],">")

group_probe_detection <- rowSums(detection_matrix) >= (probe_det/100)*n_samples_in_group

group_probe_detection_nuID <- rownames( gx[group_probe_detection==TRUE, ])

cat(" Number of probes in group [",n,"] with signal intensity greater than the mean intensity of the negative control beads = ", length(group_probe_detection_nuID) ,"\r","\n")

cat(" Writing probe list to ",paste(out_dir,"/",project_name,".GROUP.",group_label,".detected_probes_nuID.txt",sep=""), "\r","\n" )

det_probes <- as.data.frame(group_probe_detection_nuID)

colnames(det_probes) <- c("nuID")

write.table( det_probes ,file=paste(out_dir,"/",project_name,".GROUP.",group_label,".detected_probes_nuID.txt",sep=""),row.names=FALSE,quote=FALSE,col.names=FALSE)

}

####################################
## Y CHROM EXPRESSION IN XIST MALES
#####################################
cat(" Y Chromosome probe detection based on XIST MALEs","\r","\n")
xist_MALES <- pData(eset_bg)$xist_gender=="MALE"
gx_y <- exprs(eset_bg[fData(eset_bg)$CHR=="Y",])
detection_matrix_y <- sweep( gx_y[,xist_MALES],1, neg_mean[xist_MALES ],">")
y_probe_detection <- rowSums(detection_matrix_y) >= (probe_det/100) * sum(xist_MALES==TRUE)
y_probe_detection_nuID <- rownames( gx_y[y_probe_detection, ])
y_det_probes <- as.data.frame(y_probe_detection_nuID)
colnames(y_det_probes) <- c("nuID")
write.table( y_det_probes ,file=paste(out_dir,"/",project_name,".GROUP.Y.detected_probes_nuID.txt",sep=""),row.names=FALSE,quote=FALSE,col.names=FALSE)

##################################
## writing final good probe list
##################################

cat(" writing final good probe list to ",paste(out_dir,"/",project_name,".detected_probes_nuID_final.txt",sep=""),"\r","\n")
system(paste("cat ",out_dir,"/",project_name,"****.detected_probes_nuID.txt | sort | uniq >> ",out_dir,"/",project_name,".detected_probes_nuID_final.txt", sep="") )
good_probes <- read.table(file=paste(out_dir,"/",project_name,".detected_probes_nuID_final.txt",sep=""), head=FALSE)
write.table( good_probes,file=paste(out_dir,"/",project_name,".good_probes.txt",sep=""),)
good_probes <- paste(good_probes[,1])
n_good_probes <- length(good_probes)
cat(" Total number of good probes = ",n_good_probes,"\n","\r" )
good_probes_annotation <- fData(eset_bg[paste(good_probes,sep=""),])
head(good_probes_annotation)


########################################
# add $good_probe annotation to eset_bg
#########################################
fData(eset_bg)$good_probe <- fData(eset_bg)$nuID %in% good_probes
save(eset_bg, file=paste(out_dir,"/",project_name,".eset_bg.RData",sep=""))
write_expression_files(eset=eset_bg,outfile=paste(out_dir,"/",project_name,".eset_bg",sep="") )

# Saving good probe annotations
cat(" saving good probe annotations to ",paste(out_dir,"/",project_name,".detected_probes_nuID_final.***",sep=""),"\r","\n")
save(good_probes_annotation, file=paste(out_dir,"/",project_name,".detected_probes_nuID_final.RData",sep="") )
write.table(good_probes_annotation, file=paste(out_dir,"/",project_name,".detected_probes_nuID_final.txt",sep=""),quote=F,sep="\t",row.names=F )

# looksee
table(good_probes_annotation$CHROMOSOME)

# plot
plot(good_probes_annotation$n_detected_call_rate,ylim=c(0,1),pch="*",main="Probe Call Rate: Detected Probes in 80% per group", ylab="Call Rate")
abline(h=0.5,col="grey",lty=2);abline(h=0.8,col="red")

# plot pdf
dev.off()
pdf(file=paste(out_dir,"/",project_name,".eset_bg.detected_probe_call_rate.pdf",sep=""),width=8,height=8)
plot(good_probes_annotation$n_detected_call_rate,ylim=c(0,1),pch="*",main="Probe Call Rate: Detected Probes in 80% per group", ylab="Call Rate")
abline(h=0.5,col="grey",lty=2);abline(h=0.8,col="red")
dev.off()
```

****

Transform and Normalise
=======================
See **Comparison of normalization methods for Illumina BeadChip HumanHT-12 v3**
BMC Genomics. 2010; 11: 349. Ramona Schmid et al
http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3091625&tool=pmcentrez&rendertype=abstract
Figure 10 Pearson correlation of log2 ratios for different normalization methods and qRT-PCR. 
This study selects bg_rma_log_rsn as the best.
Here we have used a better method for background correction, followed by log2 transformation and robust-splince normalisation (rsn).
The robust spline normalization (RSN) algorithm combines the features of quantile and loess normalization.

## 1. lumiExpresso
```{r lumiExpresso, eval=TRUE}
# See Comparison of normalization methods for Illumina BeadChip HumanHT-12 v3.
# BMC Genomics. 2010; 11: 349.
# Ramona Schmid et al
# Figure 10 Pearson correlation of log2 ratios for different normalization methods and qRT-PCR. 
# This study selects bg_rma_log_rsn as the best
# http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3091625&tool=pmcentrez&rendertype=abstract

# log2 > rsn
eset_bg_log2_rsn_0 <- lumiExpresso(eset_bg, bg.correct = FALSE,
                                    variance.stabilize=TRUE, 
                                    varianceStabilize.param=list(method=paste(transform_method,sep="")), 
                                    normalize.param=list(method=paste(norm_method,sep="")), 
                                    verbose=FALSE);
```

## 2. Save Transformed and Normalised data pre-sample removal
```{r save_eset_bg_log2_rsn_0, eval=TRUE}
# save log2 > rsn
save(eset_bg_log2_rsn_0, file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.RData",sep="")  , compress=T)
```

## 3. Write Expression data files for Transformed and Normalised data pre-sample removal
```{r write_expression_files_eset_bg_log2_rsn_0, eval=TRUE}
# write_expression_files
write_expression_files(eset=eset_bg_log2_rsn_0,outfile= paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0",sep="") )
```


QC Plots of `eset_bg_log2_rsn_0` pre-sample removal
-------------------------------------------------------------------

## basic_qc_plot_lumi eset_bg_log2_rsn_0 pre-sample removal
```{r basic_qc_plot_lumi_eset_bg_log2_rsn_0, fig.height=7, fig.width=10,eval=TRUE}
# basic plots
# plot to screen
basic_qc_plot_lumi(eset_bg_log2_rsn_0)
par(def.par)
```

```{r basic_qc_plot_lumi_eset_bg_log2_rsn_0_pdf,eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.basic_qc_plot_lumi.pdf",sep=""),width=11,height=8)
basic_qc_plot_lumi(eset_bg_log2_rsn_0)
dev.off()
par(def.par)
```

## coloured_dendrogram_lumi eset_bg_log2_rsn_0 pre-sample removal
```{r coloured_dendrogram_lumi_eset_bg_log2_rsn_0, fig.height=7, fig.width=10,eval=TRUE}
# coloured_dendrogram_lumi
# plot to screen
coloured_dendrogram_lumi(eset_bg_log2_rsn_0)
par(def.par)
```

```{r coloured_dendrogram_lumi_eset_bg_log2_rsn_0_pdf, eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.coloured_dendrogram_lumi.pdf",sep=""),width=11,height=8)
coloured_dendrogram_lumi(eset_bg_log2_rsn_0)
dev.off()
par(def.par)
```
 
## pca_plot_lumi eset_bg_log2_rsn_0 pre-sample removal
```{r pca_plot_lumi_eset_bg_log2_rsn_0, fig.height=7, fig.width=7,eval=TRUE}
# PCA plots
# plot to screen
pca_plot_lumi(eset_bg_log2_rsn_0)
par(def.par)
```

```{r pca_plot_lumi_eset_bg_log2_rsn_0_pdf,eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.pca_plot_lumi.pdf",sep=""),width=7,height=7)
pca_plot_lumi(eset_bg_log2_rsn_0)
dev.off()
par(def.par)
```

## SampleNetwork Plots eset_bg_log2_rsn_0 pre-sample removal
```{r sampleNetwork_plot_all_lumi_eset_bg_log2_rsn_0, fig.height=10, fig.width=10,eval=TRUE}
# SampleNetwork Plots
# plot to screen
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_0, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_0, colBy="group");
par(def.par)
```

```{r sampleNetwork_plot_all_lumi_eset_bg_log2_rsn_0_pdf,eval=TRUE}
pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.sampleNetwork_plot_all_lumi.pdf",sep=""),width=8,height=8)
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_0, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn_0, colBy="group");
dev.off()
par(def.par)
```


*****

SampleNetwork : Id outlier samples 
===================================

Adapted from : ***Network methods for describing sample relationships in genomic datasets: application to Huntington's disease. Michael C Oldham et al.***
BMC Syst Biol. 2012; 6: 63.
http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3441531/?tool=pmcentrez&report=abstract

## Basic Iterative SampleNetwork outlier removal
```{r basic_sampleNetworkIterate, eval=TRUE}
# Id outlier samples
# Adapted from : Network methods for describing sample relationships in genomic datasets: application to Huntington's disease. Michael C Oldham et al.
# BMC Syst Biol. 2012; 6: 63.
# http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3441531/?tool=pmcentrez&report=abstract
# Oginal Code : 

# basic_sampleNetworkIterate
ISAoutliers <- basic_sampleNetworkIterate(eset=eset_bg_log2_rsn_0,
                                          col_by_chip=0, 
                                          groups="byGroup" ,
                                          outfile=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0",sep=""), 
                                          IACthresh=0.95, 
                                          sd_thrs=iac_sd_thrs );

# Outliers
outlier_samples <- ISAoutliers$iac_outlier_samples
cat(" number of outlier samples =", length(outlier_samples),"/",dim(eset_bg)["Samples"],"=",length(outlier_samples)/dim(eset_bg)["Samples"],"\r","\n")

save(outlier_samples, file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.SampleNetwork_outlier_samples.RData",sep="") )

# print outlier samples
ISAoutliers$iac_outlier_samples

# read in stats and lok=oksee
outlier_stats <- read.table(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn_0.group.basic_sampleNetworkIterate_summary.csv",sep=""),sep=",",head=T)
outlier_stats[,1:8]

# get pdata for outliers
SampleNetWork_outliers <- pData(eset_bg_log2_rsn_0[,outlier_samples])
n_unique_study_id_outliers <- length(unique(SampleNetWork_outliers$Study_ID))

# flag and save
pData(eset_bg_log2_rsn_0)$SampleNetWork_outlier <- pData(eset_bg_log2_rsn_0)$sampleID %in% ISAoutliers$iac_outlier_samples
save(eset_bg_log2_rsn_0, file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.RData",sep="")  , compress=T)
```

*****

Remove SampleNetWork outlier samples 
=======================================
```{r remove_SampleNetwork_outliers, eval=TRUE}
# remove outlier_samples
eset_bg_log2_rsn <- removeSamples_eset_lumi(eset=eset_bg_log2_rsn_0,sampleRemove=outlier_samples)
eset_bg_log2_rsn
str(eset_bg_log2_rsn)
# save 
save(eset_bg_log2_rsn, file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.RData",sep="")  , compress=T)
```

```{r write_expression_files_eset_bg_log2_rsn_post_sampleNetwork, eval=TRUE}
# write_expression_files
write_expression_files(eset=eset_bg_log2_rsn,outfile= paste(out_dir,"/",project_name,".eset_bg_log2_rsn",sep="") )
```

QC Plots of Transformed and Normalised data after outlier removal `eset_bg_log2_rsn`
------------------------------------------------------------------------------------

## basic_qc_plot_lumi eset_bg_log2_rsn
```{r basic_qc_plot_lumi_eset_bg_log2_rsn, fig.height=7, fig.width=10,eval=TRUE}
# basic plots
# plot to screen
basic_qc_plot_lumi(eset_bg_log2_rsn)
par(def.par)

pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.basic_qc_plot_lumi.pdf",sep=""),width=11,height=8)
basic_qc_plot_lumi(eset_bg_log2_rsn)
dev.off()
par(def.par)
```

## coloured_dendrogram_lumi eset_bg_log2_rsn
```{r coloured_dendrogram_lumi_eset_bg_log2_rsn, fig.height=7, fig.width=10,eval=TRUE}
# coloured_dendrogram_lumi
# plot to screen
coloured_dendrogram_lumi(eset_bg_log2_rsn)
par(def.par)

pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.coloured_dendrogram_lumi.pdf",sep=""),width=11,height=8)
coloured_dendrogram_lumi(eset_bg_log2_rsn)
dev.off()
par(def.par)

```

## pca_plot_lumi eset_bg_log2_rsn
```{r pca_plot_lumi_eset_bg_log2_rsn, fig.height=7, fig.width=7,eval=TRUE}
# PCA plots
# plot to screen
pca_plot_lumi(eset_bg_log2_rsn)
par(def.par)

pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.pca_plot_lumi.pdf",sep=""),width=7,height=7)
pca_plot_lumi(eset_bg_log2_rsn)
dev.off()
par(def.par)
```

## SampleNetwork Plots eset_bg_log2_rsn
```{r sampleNetwork_plot_all_lumi_eset_bg_log2_rsn, fig.height=10, fig.width=10,eval=TRUE}
# SampleNetwork Plots
# plot to screen
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn, colBy="group");
par(def.par)

pdf(file=paste(out_dir,"/",project_name,".eset_bg_log2_rsn.sampleNetwork_plot_all_lumi.pdf",sep=""),width=8,height=8)
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn, colBy="chip");
sampleNetwork_plot_all_lumi(eset_bg_log2_rsn, colBy="group");
dev.off()
par(def.par)
```
